\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[]{hyperref}

\title{Meeting 18/4}
\author{Project Group 26}
\date{\today}

\begin{document}

\maketitle

\section{Meeting}

Journal is updated. Larisa has added to the contract concerning absence. When scheduling meetings, it is important that Larisa is included in the discussion on time and place.

We've had a hard time with, problem statement, we have made an initial time line based in the study guide.

Problem statement is fine, why is time sync a problem? When we submit abstract, we will know how to compare the algorithms. It's still vague. Questions on how we are going to split up, Larisa questions if we will implement a algorithm. Next week we need to be specific (a concrete statement). She will compare us to the previous years group. Here three people implemented an algorithm. Don't forget include all the analysis, when we are reading.
If we don't we a plan it will be rushed in the end. Comparison can be difficult, want can we compare on. 

Thomas suggests looking at the classical algorithms Berkley etc.

We need to find a balance between "unable to do", and throwing very interesting stuff out the window.

Suggestions: the easiest way to implement will be paxos to slap something onto it, a paper from 2005 she was unable to find an implementation, so it's new. Each paper has analysis on different algorithms, so this should be included.   

Our halfway seminar, is at 12 one of the days, presentation should be ready if not polished. At the midway seminar each person should present a part (equal parts).

We need to find out what to do, it's a major task. 

\end{document}
