\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}

\title{Group Contract}
\author{Project Group 26}
\date{April 2018}

\begin{document}

\maketitle

\section{Group meetings}

The frequency of the meetings is to be 2 to 3 times a week. We will create a Facebook group, where we will arrange meetings and other details. If a person fails to cancel an appointment with the group, there will be slight punishment (maybe with cake). However the group expects communication from each individual if one cannot make it to a meeting.

At the last group meeting of each week, we will make a joint entry into the journal, summarizing the accomplishments make that given week.

\section{Meetings with supervisor}

Regarding the frequency of meetings with the supervisor we expect to meet once a week, preferably at the end of the work week.

We should all be present, and actively participating. One person is appointed to summarize the meeting, and the summary is made available to everyone involved after the meeting. Before each meeting we should discuss and plan the content and focus points of said meeting to avoid wasting time. 

\subsection{Being late or absent}
Being late to meetings is not acceptable for both the group and the supervisor. If a group member or the supervisor knows already that he/she would be late, he/she should notify other participants. Being late more than 10 minutes will be considered as an absence.

We shouldn't be absent to a meeting without prior notice and a good cause to be absent. We understand that it is better to notify in advance and reschedule than to be absent and penalized. 

According to 6.4.3 of the study guide, if we are absent for the third time, we have to provide a documented proof of our absence (e.g medical certificate). In case if we can't do this, we might not be recommended for participation in the final exam. 

\section{Ambition level}

As for ambition level we will strive towards a grade in the higher end of the spectrum, however we will also be satisfied with 10, or at worst 7.

\section{Contributions from each member}

Work is distributed as the project progresses. Deadlines are set for each part. In the case of a conflict or missed deadline, the issue is resolved within the group on a case-by-case basis.

\section{Conflicts}
In case of any conflicts among the team members that can not be resolved internally, we will speak to our supervisor. In case of a conflict between the group and the supervisor, we will speak to the course instructor together with the supervisor.
\end{document}
